Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-07-12 doc/manuscript_docx_copy.Rmd
Added steps to make a .docx copy of the manuscript document. Also added to the supplementary table export step to export .docx copies of the supplementary data with some added formatting.

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,\ddagger,*
    orcid: 0000-0002-3477-6291
affiliation:
  - num: 1
    address: |
      Department of Crop and Soil Sciences, NCSU, Raleigh, NC, USA
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
 In current markets, the primary uses for soybean seed are in products derived from it’s oil or protein content. However, growers are compensated based on seed yield  so a more valuable crop is the one that is high yielding with an optimum combination of protein and oil. A negative correlation of seed protein with seed yield and oil makes simultaneous improvement of  these traits difficult but not impossible through conventional breeding. A selection of lines with exceptional yield and seed composition qualities was made from two recombinant inbred line (RIL) soybean mapping populations to identify high protein and/or high oil lines with yield comparable to elite cultivars.  The performance of these RILs were evaluated in multiple environments and several genotypes were identified with yield comparable to that of high yielding check cultivars with seed protein and/or oil superior to the checks. These genotypes will provide breeders with additional sources of germplasm for continuing efforts to improve seed composition traits without compromising seed yield, and provide growers with more profitable varieties.
keywords: |
  yield; protein; oil;soybean; protein meal
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  The authors declare no conflict of 
  interest.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: RIL
    long: Recombinant inbred line
  - short: RCBD
    long: Randomized complete block design
  - short: CLA
    long: Central crops research station
  - short: CAS
    long: Caswell research farm
  - short: PLY  
    long: Tidewater research station
output: rticles::mdpi_article
---

@@

@@

The agronomic traits recorded in the field were height, lodging, maturity date, and a composite agronomic score. Lodging was scored on a scale of 1-5 where 5 indicates that all plants in a plot are on the ground, and a score of 1 indicates that all plants are erect[@fehr1987soybeans]. The agronomic score aimed to capture other traits of value such as visual estimation of pod load and plot uniformity to provide a general score of a line’s agronomic desirability. Agronomic score was recorded on a scale of 1-5 as well, with 1 identifying the best lines of a population, and 5 the worst. Maturity was recorded at the R8 maturity date and was recorded as the number of days after September 1. Height was measured in inches from the soil to the top of the plant.

@@

## Phenotypic correlations
A strong negative correlation (r = -0.93) was observed in both tests between seed protein and seed oil content. A weak negative correlation (r = -0.29) was observed between seed protein and yield in both tests which was statistically insignificant in either population. Weak positive correlations were also observed between seed oil and seed yield in both populations. A correlation coefficient of r = 0.24 was observed in Test 1 while a coefficient of r = 0.28 was observed in Test 2. These correlation coefficients were not statistically significant in either population. The pairwise correlation coefficients for Tests 1 and 2 are presented in supplementary figures 1 and 2, respectively.

@@

# Get the link to the lsmean scatterplot
mean_plot_link <- phenotype_summary_plots[grepl("lsmean_scatterplot", phenotype_summary_plots)]

@@

```{=tex}
\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{`r mean_plot_link`}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}
```

@@

## Yield contrasts
Several RILs in each of the tests had yield that was comparable to that of their check cultivars as per contrasts between RIL means and the average of the check cultivars for each test. No genotypes had yield that was higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content. Such genotypes are extremely valuable for production of soybean meal with higher protein. As mentioned earlier, the meal protein of current U.S. cultivars is mostly below the minimum market standard of 47.5%. The meal protein is calculated by the following formula that takes both seed protein and oil contents into account:

@@

The meal protein contents of these lines with high protein and normal oil contents ranged from 103.8% – 108.5% compared to the meal protein contents of the checks. Breeding line N18-1783 from test 2 had the maximum protein meal from among these lines. (Supplementary table 2).

@@

## Genotypes with comparable seed yield and seed oil and superior seed protein content
Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2.  Summary data for these lines is given in table 1.

```{r elite_geno_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$elite_genos)

@@

## Genotypes with comparable yield and superior protein
Seventeen RILs had a similar yield and superior protein content to the check cultivars. Summary data for the seed protein and seed yield of these lines is given in table 2. From among these RILs, N18-1627 had the highest seed yield relative to the checks included in the test with an average yield that was 105.9% that of the check average of Test 2. Genotype N18-1763 had the highest seed protein content on average with a protein content that was 116.8% that of the check average of Test 1. We observed a negative correlation between seed protein and seed yield in both tests so RILs with seed protein content that was above the check average tended to have a seed yield content that was comparatively lower.

```{r high_po_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$hp_sy_table)

>>>>

---
title: "Identification of high-yielding soybean lines with exceptional seed composition qualities"
author: "Jay Gillenwater"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
bibliography: ./manuscript/mybibfile.bib  
---

@@

# Abstract

 In current markets, the primary uses for soybean seed are in products derived from it’s oil or protein content. However, growers are compensated based on seed yield  so a more valuable crop is the one that is high yielding with an optimum combination of protein and oil. A negative correlation of seed protein with seed yield and oil makes simultaneous improvement of  these traits difficult but not impossible through conventional breeding. A selection of lines with exceptional yield and seed composition qualities was made from two recombinant inbred line (RIL) soybean mapping populations to identify high protein and/or high oil lines with yield comparable to elite cultivars.  The performance of these RILs were evaluated in multiple environments and several genotypes were identified with yield comparable to that of high yielding check cultivars with seed protein and/or oil superior to the checks. These genotypes will provide breeders with additional sources of germplasm for continuing efforts to improve seed composition traits without compromising seed yield, and provide growers with more profitable varieties.

@@

The agronomic traits recorded in the field were height, lodging, maturity date, and a composite agronomic score. Lodging was scored on a scale of 1-5 where 5 indicates that all plants in a plot are on the ground, and a score of 1 indicates that all plants are erect[@fehr1987soybeans]. The agronomic score aimed to capture other traits of value such as visual estimation of pod load and plot uniformity to provide a general score of a line's agronomic desirability. Agronomic score was recorded on a scale of 1-5 as well, with 1 identifying the best lines of a population, and 5 the worst. Maturity was recorded at the R8 maturity date and was recorded as the number of days after September 1. Height was measured in inches from the soil to the top of the plant.

@@

## Phenotypic correlations

A strong negative correlation (r = -0.93) was observed in both tests between seed protein and seed oil content. A weak negative correlation (r = -0.29) was observed between seed protein and yield in both tests which was statistically insignificant in either population. Weak positive correlations were also observed between seed oil and seed yield in both populations. A correlation coefficient of r = 0.24 was observed in Test 1 while a coefficient of r = 0.28 was observed in Test 2. These correlation coefficients were not statistically significant in either population. The pairwise correlation coefficients for Tests 1 and 2 are presented in supplementary figures 1 and 2, respectively.

@@

# Get the link to the lsmean scatterplot
mean_plot_link <- phenotype_summary_plots[grepl("lsmean_scatterplot.png", phenotype_summary_plots)]

@@

![**Figure 1**: Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (**A**) and Test 2 (**B**). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.](`r mean_plot_link`)

@@

## Yield contrasts

Several RILs in each of the tests had yield that was comparable to that of their check cultivars as per contrasts between RIL means and the average of the check cultivars for each test. No genotypes had yield that was higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content. Such genotypes are extremely valuable for production of soybean meal with higher protein. As mentioned earlier, the meal protein of current U.S. cultivars is mostly below the minimum market standard of 47.5%. The meal protein is calculated by the following formula that takes both seed protein and oil contents into account:

@@

The meal protein contents of these lines with high protein and normal oil contents ranged from 103.8% -- 108.5% compared to the meal protein contents of the checks. Breeding line N18-1783 from test 2 had the maximum protein meal from among these lines. (Supplementary table 2).

@@

## Genotypes with comparable seed yield and seed oil and superior seed protein content

Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2. Summary data for these lines is given in table 1.

```{r elite_geno_table, echo = FALSE}

knitr::knit_print(elite_genotype_tables$elite_genos$word_table)

@@

## Genotypes with comparable yield and superior protein

Seventeen RILs had a similar yield and superior protein content to the check cultivars. Summary data for the seed protein and seed yield of these lines is given in table 2. From among these RILs, N18-1627 had the highest seed yield relative to the checks included in the test with an average yield that was 105.9% that of the check average of Test 2. Genotype N18-1763 had the highest seed protein content on average with a protein content that was 116.8% that of the check average of Test 1. We observed a negative correlation between seed protein and seed yield in both tests so RILs with seed protein content that was above the check average tended to have a seed yield content that was comparatively lower.

```{r high_po_table, echo = FALSE}

knitr::knit_print(elite_genotype_tables$hp_sy_table$word_table)
